---
title: "Chapter 5(Edition Eight): 5.4, 5.5, 5.7, 5.8, 5.22(Hand), 5.23(Hand), 5.26"
author: "Jeremy Ling & Emmanuel Mejia"
date: "April 10, 2018"
output: pdf_document
---

##5.4

\textbf{An engineer suspects that the surface finish of a metal
part is influenced by the feed rate and the depth of cut. He
selects three feed rates and four depths of cut. He then conducts
a factorial experiment and obtains the following data}

\textbf{(a) Analyze the data and draw conclusions. Use $\alpha$ = 0.05.}

```{r}
library(readxl)
#surface <- read_excel("C:/Users/Emmanuel/Desktop/SPRING 2018/MATH 447/Chapter 5/surface.xlsx")

# surface wide format
surface.w <- as.data.frame(x = matrix(data = c(rep(c(.2, .25, .3), each = 3), c(74, 64, 60, 92, 86, 88, 99, 98, 102), c(79, 68, 73, 98, 104, 88, 104, 99, 95), c(82, 88, 92, 99, 108, 95, 108, 110, 99), c(99, 104, 96, 104, 110, 99, 114, 111, 107)), nrow = 9, byrow = FALSE))

names(surface.w) <- c("Feed.Rate", ".15", ".18", ".2", ".25")

# surface long format
surface <- melt(data = surface.w, id.vars = c("Feed.Rate"))
names(surface) <- c("Feed.Rate", "Depth", "Surface")
surface

####
surface.aov = aov(Surface ~ factor(Feed.Rate)*factor(Depth), surface)
summary(surface.aov)
```

$H_0: \tau_1 = ... = \tau_a = 0$

$H_1: \tau_i \ne 0$ at least one

$H_0: \beta_1 = ... = \beta_b = 0$

$H_1: \beta_j \ne 0$ at least one

$H_0: (\tau\beta)_{1 1} = ... = (\tau\beta)_{a b} = 0$

$H_1: (\tau\beta)_{ij} \ne 0$ at least one

We analyze the anova and we can see from our p-values that they are all too small. We can state the row, column, and interaction are all significantly different and we can state that this model is significant.

\textbf{(b) Prepare appropriate residual plots and comment on the
model???s adequacy.}

```{r}
surface.lm=lm(Surface ~ Feed.Rate*Depth,surface)
res=surface$Surface-fitted(surface.lm)
library(car)
qqPlot(res)
plot(fitted(surface.lm), res) 
plot(as.numeric(surface$Feed.Rate), res)
plot(as.numeric(surface$Depth), res)
```

We observe our plots and see normality is good and residuals are patternless, our model is good..

\textbf{(c) Obtain point estimates of the mean surface finish at
each feed rate.}

```{r}
total1 = (74+64+60+79+68+73+82+88+92+99+104+96)
estimate0.20 = total1/12; estimate0.20
total2 = (92+86+88+98+104+88+99+108+95+104+110+99)
estimate0.25 = total2/12; estimate0.25
total3 = (99+98+102+104+99+95+108+110+99+114+111+107)
estimate0.30 = total3/12; estimate0.30
```

\textbf{(d) Find P-values for the tests in part (a)}

```{r}
surface.aov = aov(Surface ~ Feed.Rate*Depth, surface)
summary(surface.aov)
```

The P-values are presented on the ANOVA output. The same p-values were used to determine our test in part (a). P-values are listed: 1.40e-10, 9.22e-09, 0.00156.

##5.5

\textbf{For the data in Problem 5.4, compute a 95 percent
confidence interval estimate of the mean difference in
response for feed rates of 0.20 and 0.25 in/min.}

```{r}
estimate0.20
estimate0.25
alpha = 0.05
a=3
b=4
n=3
MSe=summary(surface.aov)[[1]][4,3]
tk.cri=qtukey(.95,a,a*b*(n-1))*sqrt(2*MSe/n)
diff = estimate0.20-estimate0.25
me=c(-1,1)*tk.cri+diff
me
```

We are 95% confident that the difference between feed rates 0.20 and 0.25 is (-32.9, 0.953).

##5.7

\textbf{Johnson and Leone (Statistics and Experimental
Design in Engineering and the Physical Sciences, Wiley,
1977) describe an experiment to investigate warping of
copper plates. The two factors studied were the temperature
and the copper content of the plates. The response variable
was a measure of the amount of warping. The data were as
follows:}

\textbf{(a) Is there any indication that either factor affects the
amount of warping? Is there any interaction between
the factors? Use $\alpha$ =  0.05.}

```{r}
library(readxl)
warpCop <- read_excel("warpCop.xlsx")
warpCop$Temperature=as.factor(warpCop$Temperature) #factor A; a levels
warpCop$Copper.Content=as.factor(warpCop$Copper.Content) #factor B; b levels
warpCop.aov = aov(Warping ~ Temperature*Copper.Content, warpCop)
summary(warpCop.aov)

MSe=summary(warpCop.aov)[[1]][4,3]
```

$H_0: \tau_1 = ... = \tau_a = 0$

$H_1: \tau_i \ne 0$ at least one

$H_0: \beta_1 = ... = \beta_b = 0$

$H_1: \beta_j \ne 0$ at least one

$H_0: (\tau\beta)_{1 1} = ... = (\tau\beta)_{a b} = 0$

$H_1: (\tau\beta)_{ij} \ne 0$ at least one

After reviewing the p-values from our ANOVA we can see that temperature and copper content have small p-values and we can conclude that both factors affect the amount of warping. However, the interaction variable has a large p-value, we fail to reject and conclude no interaction has an affect on warping.

\textbf{(b) Analyze the residuals from this experiment.}

```{r}
warpCop.lm=lm(Warping ~ Temperature*Copper.Content,warpCop)
res=warpCop$Warping-fitted(warpCop.lm)
library(car)
qqPlot(res)
plot(fitted(warpCop.lm), res) 
plot(as.numeric(warpCop$Temperature), res)
plot(as.numeric(warpCop$Copper.Content), res)
```

Observing our plots, normality is good, residual plots are random and patternless. We can state that our model is good.

\textbf{(c) Plot the average warping at each level of copper content
and compare them to an appropriately scaled t distribution.
Describe the differences in the effects of the
different levels of copper content on warping. If low
warping is desirable, what level of copper content
would you specify?}

